Cross-sectoral conservation of the Indian Ocean

### 2022-02-15 Msc_Thesis_Lea_v2.Rmd
Update Msc_Thesis_Lea_v2.Rmd

+ how to group all files (ex: all EBSAs + IBAs + IMMAs + vents + seamounts...) into one, to create ConsFeatures?
+ should we use Sharma (2011) + CRU Consulting (2019) to assign USD values to PMN areas? see PMN subsection
+ shipping raster (Halpern et al., 2013) is too heavy to be converted to polygon, what coupd be other methods (not using as.polygon)?
+ shipping raster : cells have certain values (# of ships recorded??I will check for Tuesday), could allow to infer USD value of route (# of ships * approx cost of fuel for average cargo to go through 1 PU, which is 1000sqkm)? So we wouldn't try to compute the cost of deviating from any given route, but rather compute the price for each ship to go through 1 PU * the number of ships going through that PU. It sounds fairly logical to me, but maybe it's not?
+ fishing cost layer : there is a ready-made fx in SpatPlan repo (thanks Jase & Tin), should we use it for this project? see fishing subsection (but issues with plotting  as you can see)

@@

+ to reach the global target of 30% marine protection, ABNJs are optimal (largest area; common good: higher incentive to protect since lower opportunity cost for individual countries)

@@

```{r, cache = FALSE, echo = FALSE, results = "hide", message = FALSE}
#Open all the packages and get extras
options(repos = list(CRAN="http://cran.rstudio.com/")) 
library(tidyverse)
library(sf)
library(patchwork)
library(terra)
library(stars)
library(rnaturalearth)
library(rnaturalearthdata)
library(prioritizr)
library(knitr)

@@

still need to

+ **select specific fishing rasters (bycatch/demersal...)/ use the fishing cost layer  already created in Spatial Planning repo**
+ and **select specific benthic shps (seamounts, ridges, plateaus, vents?)**
+ **need to create subset without inactive vents** *filter fx not working because object=character, will ask on Slack*

@@

Input type   |  Topic                 | Reference/source (format)
:-----------:|:----------------------:|:----------------------------------:
 COST        |  Expl + Res (mining)   |  ISA (shp)
 COST        |  PMN & CFC (mining)    |  Petersen et al., 2016 (shp)
 COST        |  Shipping              |  Halpern et al., 2013 (tif)
 COST        |  Fishing               |  Halpern et al., 2013 (tif)
 CONS        |  IBAs                  |  Birdlife datazone (shp)
 CONS        |  IMMAs                 |  importantmarinemammalhabitat (shp)
 CONS        |  EBSAs                 |  CBD (shp)
 CONS        |  Benthic features      |  Blue Habitats (shp)
 CONS        |  Longhurst provinces ? |  Marine Regions (shp)
 CONS        |  Vents                 |  Beaulieu & Szanfranski (2020) (csv)

@@

Boundary
```{r PR, echo=FALSE}
ggplot(data = Bndry) + geom_sf()
```

@@

+ Important Bird Areas
+ Important Marine Mammal Areas
+ EBSAs
+ Deepsea features: vents, seamounts, ridges, plateaus? *(need to define)*

@@

**Still unable to add legend in plots so showing the code instead, but will ask on Slack**

Seamounts, ridges, plateaus, vents

+ *need to add chunck for ridges & plateaus if we decide to add them*

+ *need to check how to 'filter' out inactive vents and to cross vents (point data) with PUs and how to add in plot*

@@

Ridges/Plateaus: >30%

@@

Vents <- read.csv("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/vents.csv")
VentsActive <- filter(Vents, Activity == "active, confirmed" | Activity == "active, inferred")
PUxVents <- fSpatPlan_Get_Polyg(VentsActive, PUs)

@@

IMMA <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/IMMAs/iucn-imma.shp")
IBA <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/IBAs/MarineIBA_IndianOcean.shp")
NEIO_09<- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/NEIO_9_EBSA/NEIO_9_EBSA.shp")
NWIO_14 <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/NWIO_14_EBSA/NWIO_14_EBSA.shp")
SIO_11 <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_11_EBSA-GIS shapefile/SIO_11_EBSA.shp")
SIO_19 <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_19_EBSA-GIS%20shapefile/SIO_19_EBSA.shp")
SIO_22 <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_22_EBSA-GIS shapefile/SIO_22_EBSA.shp")
SIO_23 <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_23_EBSA-GIS shapefile/SIO_23_EBSA.shp")
SIO_30 <- st_read  ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_30_EBSA-GIS shapefile/SIO_30_EBSA.shp")
SIO_32 <- st_read  ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_32/SIO_32_EBSA.shp")
SIO_35<- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_35_EBSA-GIS shapefile/SIO_35_EBSA.shp")
SIO_36 <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_36_EBSA-GIS shapefile/SIO_36_EBSA.shp")
SIO_37  <-st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_37_EBSA-GIS shapefile/SIO_37_EBSA.shp")
Seamounts <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/BlueHabs/Seamounts.shp")
```

```{r Conservation features plot, echo = TRUE, message = FALSE, warning = FALSE}

ggplot() + 
  geom_sf(data = world) +
  geom_sf(data = IBA, size = 1, color = "green", fill = "green") + 
  geom_sf(data = IMMA, size = 1, color = "blue", fill = "blue")+ 
  geom_sf(data = NEIO_09, size = 1, color = "grey", fill = "grey")+ 
  geom_sf(data = NWIO_14, size = 1, color = "grey", fill = "grey")+
  geom_sf(data = SIO_11, size = 1, color = "grey", fill = "grey")+
  geom_sf(data = SIO_19, size = 1, color = "grey", fill = "grey")+
  geom_sf(data = SIO_22, size = 1, color = "grey", fill = "grey")+
  geom_sf(data = SIO_23, size = 1, color = "grey", fill = "grey")+
  geom_sf(data = SIO_30, size = 1, color = "grey", fill = "grey")+
  geom_sf(data = SIO_32, size = 1, color = "grey", fill = "grey")+
  geom_sf(data = SIO_35, size = 1, color = "grey", fill = "grey")+
  geom_sf(data = SIO_36, size = 1, color = "grey", fill = "grey")+
  geom_sf(data = SIO_37, size = 1, color = "grey", fill = "grey")+
  geom_sf(data = Seamounts, size = 1, color = "black", fill = "black")+
  ggtitle("IBA, IMMA, EBSAs, Seamounts") + 
  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)
```

## Cost layer

@@

*Still unable to add legend so showing the code, will ask on Slack*

@@

```{r Mining layer plot, echo = TRUE, message = FALSE, warning = FALSE}
ggplot() + 
  geom_sf(data = world) +
  geom_sf(data = CFC, size = 1, color = "orange", fill = "orange") + 
  geom_sf(data = PMN, size = 1, color = "brown", fill = "brown") + 
  geom_sf(data = Expl_PMN, size = 1, color = "red", fill = "red") + 
  geom_sf(data = Expl_PMS, size = 1, color = "red", fill = "red") +
  geom_sf(data = Expl_CFC, size = 1, color = "red", fill = "red") +
  geom_sf(data = Res_PMN, size = 1, color = "red", fill = "red") +
  geom_sf(data = Res_CFC, size = 1, color = "red", fill = "red") +
  ggtitle("PMN, CFC, Exploration & Reserved areas") + 
  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)

## Shipping

>>>>

+ how to group all files (ex: all EBSAs + IBAs + IMMAs + vents + seamounts...) into one, to create ConsFeatures? Or is it enough to have columns next to PUs? RESOLVED: yes, stacking = creating columns next to PUs.
+ should we use Sharma (2011) + CRU Consulting (2019) to assign USD values to PMN areas? see PMN subsection. RESOLVED, yes, can use for this project (need to cross-check if want to publish)
+ shipping raster (Halpern et al., 2013) is too heavy to be converted to polygon, what could be other methods (not using as.polygon)? RESAMPLE if resolution to fine
+ shipping raster : cells have certain values (# of ships recorded??I can't figure from the attached files, metadata is confusing), could allow to infer USD value of route (# of ships * approx cost of fuel for average cargo to go through 1 PU, which is 1000sqkm)? So we wouldn't try to compute the cost of deviating from any given route, but rather compute the price for each ship to go through 1 PU * the number of ships going through that PU. It sounds fairly logical to me, but maybe it's not? RESOLVED: check 2008 metada, run prio problem with only shipping
+ fishing cost layer : there is a ready-made fx in SpatPlan repo (thanks Jase & Tin), should we use it for this project? see fishing subsection (but issues with plotting  as you can see). RESOLVED: yes, use fx from Jase & Tin. Change scale to remove outliers/crop to planning region.
+ I'm using echo = FALSE, message = FALSE, warning = FALSE, results = "hide" and still can't get rid of some coding output in the .html :( 
+ I st_to_sf the vents from .csv with PUs' crs but have some issues intersecting the points with polygs, will look more into it. ASK ON SLACK

@@

+ to reach the global target of 30% marine protection, ABNJs are optimal (largest area; common good: higher incentive to protect since lower opportunity cost for individual countries)
+ but maybe we can also include non ABNJ? (because shipping data more interesting in EEZs, tbd later)

@@

```{r, echo = FALSE, results = "hide", message = FALSE}
options(repos = list(CRAN="http://cran.rstudio.com/")) 
library(tidyverse)
library(sf)
library(patchwork)
library(terra)
library(stars)
library(rnaturalearth)
library(rnaturalearthdata)
library(prioritizr)
library(knitr)

@@

+ **using the fishing cost layer  already created in Spatial Planning repo**
+ deep-sea features = seamounts, plateaus, vents

still need to intersect vent points with PUs

@@

Input type   |  Topic                 | Reference/source (format)
:-----------:|:----------------------:|:----------------------------------:
 COST        |  Expl + Res (mining)   |  ISA (shp)
 COST        |  PMN & CFC (mining)    |  Petersen et al., 2016 (shp)
 COST        |  Shipping              |  Halpern et al., 2013 (tif)
 COST        |  Fishing               |  Jase & Tin
 CONS        |  IBAs                  |  Birdlife datazone (shp)
 CONS        |  IMMAs                 |  importantmarinemammalhabitat (shp)
 CONS        |  EBSAs                 |  CBD (shp)
 CONS        |  Benthic features      |  Blue Habitats (shp)
 CONS        |  Longhurst provinces ? |  Marine Regions (shp)
 CONS        |  Vents                 |  Beaulieu & Szanfranski (2020) (csv)

@@

@@

+ Important Bird Areas
+ Important Marine Mammal Areas
+ EBSAs
+ Deepsea features: active vents, seamounts, plateaus?

@@

Seamounts,plateaus, vents

*need to cross vents with PUs*

@@

Plateaus: >30%

@@

Plateaus <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/BlueHabs/Plateaus.shp"
PUxPlateaus <- fSpatPlan_Get_Polyg(Plateaus, PUs)  
PUs <- PUs %>% 
  mutate(Plateaus = PUxPlateaus)

Vents <- read.csv("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/vents.csv")
VentsActive <- filter(Vents, Activity == "active, confirmed" | Activity == "active, inferred")
VentsActiveSF<- st_as_sf(VentsActive, coords = c("Latitude", "Longitude"), crs = st_crs(PUs))
PUxVents <- st_intersects(PUs, VentsActiveSF) %>% 
    as.logical() 
PUs <- PUs %>% 
  mutate(Vents = PUxVents) ##issue: NAs everywhere, probably because trying to cross points with polygs?
```

@@

IMMA <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/IMMAs/iucn-imma.shp")
IBA <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/IBAs/MarineIBA_IndianOcean.shp")
NEIO_09<- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/NEIO_9_EBSA/NEIO_9_EBSA.shp")
NWIO_14 <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/NWIO_14_EBSA/NWIO_14_EBSA.shp")
SIO_11 <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_11_EBSA-GIS shapefile/SIO_11_EBSA.shp")
SIO_19 <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_19_EBSA-GIS%20shapefile/SIO_19_EBSA.shp")
SIO_22 <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_22_EBSA-GIS shapefile/SIO_22_EBSA.shp")
SIO_23 <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_23_EBSA-GIS shapefile/SIO_23_EBSA.shp")
SIO_30 <- st_read  ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_30_EBSA-GIS shapefile/SIO_30_EBSA.shp")
SIO_32 <- st_read  ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_32/SIO_32_EBSA.shp")
SIO_35<- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_35_EBSA-GIS shapefile/SIO_35_EBSA.shp")
SIO_36 <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_36_EBSA-GIS shapefile/SIO_36_EBSA.shp")
SIO_37  <-st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/EBSAs/SIO_37_EBSA-GIS shapefile/SIO_37_EBSA.shp")
Seamounts <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/BlueHabs/Seamounts.shp")
Plateaus <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Biodiv_features/BlueHabs/Plateaus.shp")
```

Conservation features

```{r Conservation features plot, echo = FALSE, message = FALSE, warning = FALSE}

ggplot() + 
  geom_sf(data = world) +
  geom_sf(data = NEIO_09,  aes(color="A"), show.legend = "line")+ 
  geom_sf(data = NWIO_14,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_11,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_19,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_22,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_23,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_30,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_32,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_35,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_36, aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_37,  aes(color="A"), show.legend = "line")+
  geom_sf(data = Seamounts,  aes(color="B"), show.legend = "line")+
  geom_sf(data = Plateaus,  aes(color="C"), show.legend = "line")+
  geom_sf(data = IBA,  aes(color="D"), show.legend = "line") + 
  geom_sf(data = IMMA,  aes(color="E"), show.legend = "line")+ 
  geom_sf(data = VentsActiveSF,  aes(color="F"), show.legend = "line")+ 
  ggtitle("IBA, IMMA, EBSAs, Seamounts, Plateaus, Vents")+ 
  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE) +
  scale_color_manual(values = c("A" = "grey", "B" = "black", "C" = "purple", "D" = "green", "E" = "blue", "F" = "red"), 
                     labels = c("EBSAs", "Seamounts", "Plateaus", "IBAs", "IMMAs", "Active vents"),
                     name = "Legend")

```
## Cost layer

@@

@@

Mining areas

```{r Mining layer plot, echo = FALSE, message = FALSE, warning = FALSE}
ggplot() + 
  geom_sf(data = world) +
  geom_sf(data = CFC, aes(color="A"), show.legend = "line") + 
  geom_sf(data = PMN, aes(color="B"), show.legend = "line") + 
  geom_sf(data = Expl_PMN, aes(color="C"), show.legend = "line") + 
  geom_sf(data = Expl_PMS, aes(color="C"), show.legend = "line") +
  geom_sf(data = Expl_CFC, aes(color="C"), show.legend = "line") +
  geom_sf(data = Res_PMN, aes(color="D"), show.legend = "line") +
  geom_sf(data = Res_CFC, aes(color="D"), show.legend = "line") +
  scale_color_manual(values = c("A" = "orange", "B" = "brown", "C" = "red", "D" = "red"), 
                     labels = c("CFC", "PMN", "Exploration areas", "Reserved areas"),
                     name = "Legend") +
  ggtitle("PMN, CFC, Exploration & Reserved areas") + 
  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)

```
## Shipping
